---
title: "ASSIGNMENT"
author: "ELLEN MASEKO"
date: "18/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message=FALSE, echo=FALSE}
#import libraries
library(tidyverse)
library(ggplot2)
library(readxl)
```


## 1. BASICS
### 1.1. Variables in R

What is the value of " B " after running these command?
```{r, }
a <- 5.0
c <- 10
t <- 12
B <- a+c/t
```
The value of B is : `r round(B,2)`


What type of vector do you get in each of the following?
```{r}
vec_1 <- c(44, 5, 0, "k")
vec_2 <- c(1, 2, 3, FALSE)
vec_3 <- c("a", "b", "c", TRUE)
vec_4 <- c(1, 2, 3, "4")
```
In vec_1 we have numeric vector

vec_2 has numeric vector

vec_3 has character vector

vec_4 has numeric vector



### 1.2. Conditional Subsetting
```{r}
vec <- c(141.2741, 141.8392, 140.5730,NA, 141.3571, 139.8689, 
         137.9384, NA,139.6620, 138.2521, 138.8635, 141.3885, NA)
```

1.2.1. Create a vector vec_1 with no missing values using conditional subsetting and explain your code
```{r}
vec_1 <- vec[!is.na(vec)]
print(vec_1)
```
##### I created vector 1 to remove all the missing values and print out only the available values. !is.na is used to only select and print out the values the are not missing.



1.2.2. Create a vector of indexes of missing values using conditional subsetting and explain your code 
```{r}
vec_2 <- vec[is.na(vec)]
print(vec_2)
```


1.2.3. Write a code to determine how many missing values are in vec.
```{r}
vec <- c(141.2741, 141.8392, 140.5730,NA, 141.3571, 139.8689, 
         137.9384, NA,139.6620, 138.2521, 138.8635, 141.3885, NA)

length(vec[is.na(vec)])
```


1.2.4. Write a R code to create a data frame which contains details of 3 students 
```{r}
name <- c("TAKUNDA", "TAKUDZWA", "KATIE")
surname <- c("NYENYA", "MACHIDA", "DU PLOOY")
age <- c(21, 20, 23)
gender <- c("female", "male", "female")

student.data <- data.frame(name, surname, age, gender)

print(student.data)
```

## 2. Functions in R

2.1. Create a function which returns a mean of an input vector and explain your code
```{r}
mean <- function(vector){
  b <- mean(vector,na.rm=TRUE)
  return(b)
}
```
I created a function which returns the mean of an input vector, I assigned the mean of the vector to "b" and return "b" and used na.rm to remove all the NA values not computing them in the mean if there are any in the vector.


2.2. Create a function which take no inputs , it reads a Titanic .csv file and returns its content and explain your code.
```{r}
df <- read.csv("train.csv")
print(df)
```
I created a df function which reads the train csv file and return all the information available in the file about the titanic (all the people's information).


## 3. Conditionals and Loops

3.1. Write a function that take an input below and returns a transformed vector below into 1 when its element is male and 2 when its female using ifelse. [8]
```{r}
g <- c("MALE","FEMALE","FEMALE","MALE","FEMALE","MALE","MALE","FEMALE", "MALE","MALE","FEMALE","MALE","MALE",
"FEMALE","MALE","MALE","MALE")
Gender<-function(){
  y<-ifelse(g == "MALE",1,2)
  return(y)
}
Gender()

```


3.2. Generate x from normal distribution and sample with replacement "M" and "F" for male and female respectively by running the code below.
```{r}
x <- rnorm(1000, 120, 5)
y <- sample(c("M","F"),size = 1000, replace = TRUE)

```


3.2.1. Use the "cbind " function in R to create one dataset and assign it to df. [2]
```{r}
df <- cbind(x, y)
print(df)
```


3.2.2. Convert df from 3.2.1 into a data.frame [2]
```{r}
df <- data.frame(df)
print(df)
```


3.2.3. Write a code using tidyverse to compute a number of males ("M") and females ("F") in 3.2.1 [3]
```{r}

x <- rnorm(1000, 120, 5)
y <- sample(c("M","F"),size = 1000, replace = TRUE)

df %>% group_by(y) %>% count()
```


3.2.4. Create a bar plot using ggplot to visualize your results in 3.2.3 [5]
```{r}
df %>% group_by(y) %>% count() %>%
  ggplot(aes(x=y, y=n))+
  geom_col()+
  labs(y="total", x="Gender", title = "Bar Graph of number of females and males")
```


3.2.5. Use mutate function to add sum of x per "M" and "F" and plot the results using sum of x per "M" and "F". [5]
```{r}
df_new <- df %>% group_by(y) %>% count() %>%
  mutate(per_no = (n/1000)*100)
ggplot(df_new, aes(y , y=per_no))+
  geom_col()
```



